Polytope notes

### 2019-12-14 notes/polytope-notes.tex
Post NeurIPS.  Took a second pass on most of current statements, but one left and want to add summary of neurips discussion later.

  It is worth noting that normal cones are generally not closed by this definition, so sometimes we want to think about the closure of the normal cone.

@@

\jessie{Not sure if we'll need this next corollary yet, but seems like it might be useful.}
\begin{corollary}[EPFL Cor 3.1.3]
  Let $P = P_1 \oplus \ldots \oplus P_n$ be a Minkowski sum of polytopes in $\reals^d$, let $F$ be a nonempty face of $P$ , and let $F_1, \ldots, F_n$ be its decomposition.
  Then $\N(F;P) = \N(F_1;P_1) \cap \ldots \cap \N(F_n; P_n)$.
\end{corollary}

@@

  \begin{definition}
    Suppose we are given a polytope $T \in \reals^d$ and set of vectors $V \in \reals^{k \times d}$.
    Call $e \in \reals^k$ the vector such that $e_i = \max_{x \in T}\inprod{v_i}{x}$.  
    We say a set of normals $V$ is \emph{complete} with respect to a polytope $T_y$ if $T_y = \{x \in \reals^d: Vx \leq E_{;y}\}$, where $E_{;y}$ is the $y^{th}$ column of $E$.

    Moreover, $V$ is complete with respect to a set of polytopes $\T$ if and only if $V$ is complete with respect to each $T \in \T$.
  \end{definition}

  \subsection{Finding $V$}
  \btw{Not sure if we want to go this far, or just argue that a complete set of normals exists.  This would be where Corollaries 1 and 2 come in though.}
  In order to find a $V$ that is complete with respect to $T$, we can look at the normal cones for each $T_i$ and find a vector in the normal cone $\N(F_y, T_y)$ in the decomposition of $F$ in $T$ for each face $F$.
  Since we know $T_1, \ldots, T_n$ are polytopes with $\H$-representations, one of the first natural steps is to find a (possibly minimal) set of valid inequalities describing each $T_i$.
  For each face $F_j$ of a given polytope $T$, we can choose a normal $v_j \in \N(F_j; T)$ and describe the face $F$ by $\{x \in T:  \inprod{v_j}{x} = \sup_{x \in T} \inprod{v_j}{x} =: e_j\}$.
  For each face $F_{jy}$ that is in the decomposition of the face $F_j$, we have $\inprod {v_j}{x} \leq e_j$ forming a tight bound on $F_j$. 
  If we do this over each $T_y$, we can call $E_y$ the vector of such $e_j$ vectors for each face of $T_y$.
  We then stack these $E_y$ vectors to form the matrix $E$.

@@

  \begin{conjecture}
  	The set of normals $V$ and vectors $E$ constructed by taking an element of the normal cone is complete for both $T$ and each $T_y \in \T$.
  \end{conjecture}

  \subsection{Once we have $V$}
  Now, for a given polytope $T(p) := \oplus p_y T_y$, we will later want to ask when a given $z \in \reals^d$ is in the polytope $T(p)$.
  First, we can claim $T(p) = \{x : Vx \leq E p \}$, where $V$ is complete for $\T$.
  To see this, first observe that $p_y T_y = \{x : Vx \leq p_y E_{;y}\}$.
  Now in taking the Minkowski sum of polytopes described by the same set of normals, we can take 
  \begin{align*}
  \oplus_y p_y T_y &= \{x_1 : Vx_1 \leq p_1 E_{;1}\} \oplus \ldots \oplus \{x_n : Vx_n \leq p_n E_{;n}\} \\
  &= \{x : Vx \leq p_1 E_{;1} + \ldots + p_n E_{;n}\}\\
  &= \{x : Vx \leq E p\}
  \end{align*}
  Observe the first to second line follows from Theorem~\ref{thm:support-minksum}.

  Now, we have $z \in T(p) \iff \inprod{v_i}{z} \leq (Ep)_i$ for all $i \in [k]$.
  We are particularly interested in when $\vec 0 \in T(p)$.  \btw{For later, common result to say $T(p) = \partial \inprod p {L(\cdot)}$, so if $0 \in T(p)$, we have a minimizer of the loss.}
  Plugging in $\vec 0$, we have $\vec 0 \in T(p)$ if and only if $E p \geq 0$

@@

@@

 	Then $\{p \in \simplex : \vec 0 \in \oplus_y p_y T_y\} = \{p \in \simplex: Ep \geq \vec 0\}$.\btw{``stuff about $E$ and $D$.''}
  \end{lemma}
  \begin{proof}
    First, let us fix a distribution $p \in \simplex$.
    We define $T(p) := \oplus_y p_y T_y$ to be the $p$-weighted Minkowski sum over the polytopes in $\T$.
    By Theorem~\ref{thm:support-minksum}, we have the support of the (weighted) Minkowski sum is the (weighted) sum of the support of each polytope.

@@

  \begin{definition}
  	\jessie{I don't like this definition, but I'm pretty sure it's correct at least.  Can we simplify it?}
  	We say a  row $X_i$ of the matrix $X$ is \emph{redundant} with respect to matrix $Y$ describing the polytope $P^Y := \{z \in A : Yz \geq \vec 0\}$ if we have $\{z \in P^Y : Xz \geq \vec 0\} = \{z \in P^Y : X^{-i}z \geq \vec 0\}$, where $X^{-i}$ is the matrix $X$ with the row $X_i$ removed.
  \end{definition}

  \begin{proposition}
  	Suppose we have polytopes $\T = \{T_1, \ldots, T_n\}$ and a set of normals $V$ that is complete for $\T$.
  	Take $E = (e_{iy})$ where $e_{iy} = \max_{x \in T_y} \inprod{v_i}{x}$, and take $D = \{p \in \simplex : Ep \geq \vec 0\}$ and $C = \{p \in A : Bp \geq \vec 0\}$.

  	Then $\{p \in \simplex : \vec 0 \in \oplus_y p_y T_y\} = C$	if and only if non-simplex rows of $B$ appear in $E$ (possibly scaled) and every other row of $E$ is redundant with respect to $B$.
  	\btw{``replace prop 1 statement with `non-simplex rows of $B$ appear in $E$ (possible scaled) and every other row of $E$ is redundant with respect to $B$.'''}
  \end{proposition}
  \begin{proof}
	\begin{itemize}
		\item [$\implies$] First, assume $C = \{p \in \simplex: \vec 0 \in \oplus_y p_y T_y\}$.
		By Proposition~\ref{prop:relate-E-B}, we know that $C = D := \{p \in \simplex : Ep \geq \vec 0\}$.
		We can re-write $D$ as $\{p \in A : [E;I_n] p \geq \vec 0\}$, building simplex constraints into the inequality rather than the domain assumption.
		By the assumed equality, every row of $B$ must then appear in $[E;I_n]$, where $I_n$ is the $n$-element identity matrix.
		As $\{p \in A : I_n p \geq \vec 0\}$ strictly defines the simplex, any non-simplex constraint of $B$ must then be described by a row of $E$, up to rescaling.

		What's more is that since the polytopes are the same, they have a unique minimal $\H$-representation (up to rescaling) since they are defined by an inequality on $\vec 0$.
		As $D$ is defined by distributions on the simplex and $C$ on the affine span of the simplex, and $\simplex \subset A$, we have the rest of the rows in $E$ necessarily being redundant with respect to $B$.
		To see this, suppose there was a row of $E$ that was not redundant with respect to $B$.
		If such a row did not partition the simplex, it would necessarily be redundant by the restriction that $C \subseteq \simplex$.
		Thus, such a row much partition the simplex, but not appear in $B$ up to rescaling.
		Again, if the constraint does not partition the cell, the constraint is redundant, and if it does, then $C \neq D$, yielding a contradiction.

		\item [$\impliedby$] Now, suppose that any non-simplex rows of $B$ appear in $E$, and every other row of $E$ is redundant with respect to $B$.
		Then we have $D = \{p\in \simplex : Ep \geq \vec 0\} = \{p \in \simplex : Bp \geq \vec 0\}$ by the assumption and this polytope is equal to $\{p \in A : Bp \geq \vec 0\} = C$ since we require $C\subseteq \simplex$, so $B$ contains any non-redundant simplex constraints.

@@

>>>>

  It is worth noting that normal cones are generally not closed by this definition, but sometimes we may want to think about the closure of the normal cone.

@@

\jessie{Not sure if we'll need this next corollary yet, but seems like it might be useful.}
\begin{corollary}[EPFL Cor 3.1.3]\label{cor:face-decomp-normal-cones}
  Let $P = P_1 \oplus \ldots \oplus P_n$ be a Minkowski sum of polytopes in $\reals^d$, let $F$ be a nonempty face of $P$ , and let $F_1, \ldots, F_n$ be its decomposition.
  Then $\N(F;P) = \N(F_1;P_1) \cap \ldots \cap \N(F_n; P_n)$.
\end{corollary}

@@

  \subsection{Notation and new definitions}
  We will use the following notation throughout the rest of this section.
  Suppose we are given a polytope $T_y \in \reals^d$ and set of vectors $V \in \reals^{k \times d}$.
  Call $e^y \in \reals^k$ the vector such that $e^y_i = \max_{x \in T_y}\inprod{v_i}{x}$.  
  For a finite set $\T = \{T_1, , \ldots, T_n\}$, let us denote the matrix $E = (e^y)_{y=1}^n$.
  \begin{definition}
    We say a set of normals $V$ is \emph{complete} with respect to a polytope $T_y$ if $T_y = \{x \in \reals^d: Vx \leq e^y\}$.

    Moreover, $V$ is complete with respect to the set of polytopes $\T$ if and only if $V$ is complete with respect to each $T_y \in \T$.
  \end{definition}

  \subsection{Finding $V$}
  \btw{Not sure if we want to go this far, or just argue that a complete set of normals exists.  This would be where Corollaries 1 and 2 come in though.}
  In order to find a $V$ that is complete with respect to $T := \oplus_y T_y$, we can look at the normal cones for each $T_i$ and find a vector in the normal cone $\N(F_y, T_y)$ in the decomposition of $F$ in $T$ for each face $F$.
  Since we know $T_1, \ldots, T_n$ are polytopes with $\H$-representations, one of the first natural steps is to find a (possibly minimal) set of valid inequalities describing each $T_i$.
  For each face $F_j$ of a given polytope $T$, we can choose a normal $v_j \in \N(F_j; T)$ and describe the face $F$ by $\{x \in T:  \inprod{v_j}{x} = \sup_{x \in T} \inprod{v_j}{x} =: e_j\}$.
  For each face $F_{jy}$ that is in the decomposition of the face $F_j$, we have $\inprod {v_j}{x} \leq e^y_j$ forming a tight bound on $F_j$. 
  If we do this over each $T_y$, we can call $e^y$ the vector of such $e^y_j$ vectors for each face of $T_y$.
  We then concatenate these $e^y$ vectors to form the matrix $E$.

@@

  \begin{conjecture}
  	The set of normals $V$ and matrix $E$ constructed by taking an element of the normal cone $\N(F, T)$ for each face $F$ in $T$ is complete for both $T$ and each $T_y \in \T$.
  \end{conjecture}
  \begin{proof}
  	Thoughts Dec 3, 2019:

  	First, to see this constructed $V$ is complete for $T$, consider that the set of normals to the facets (in the affine hull) of $T$ is complete for $T$.
  	Moreover, each additional normal is a redundant constraint, since the inner product of elements of the normal cone is less than or equal to the support for the polytope for every element of $T$.
  	Thus, this set of normals is complete for $T$.

  	To see why $V$ is complete with respect to each $T_y$, consider each case of subset inclusion.
  	First,  to see $T_y \subset \bar T_y  := \{ x : Vx \leq E_{;y}\}$, take $x \in T_y$.
  	For all $i \in [k]$, we then have $\inprod{v_i}{x} \leq H_{T_y}(v_i)$.
  	This is true be definition for all $v_i$ such that $v_i \in \N(F, T_y)$ for some $F$ a face of $T_y$.
  	We want to claim this is true for all $v_i$, i.e. for all $i$, we have $v_i \in \N(F_y, T_y)$ for some face $F_y$ of $T_y$.
  	This follows from Corollary~\ref{cor:face-decomp-normal-cones}, since we know that we must have $v_i \in \N(F, T)$ for some face $F$ of $T$, and we necessarily have $v_i \in \N(F_y, T_y)$, where $F_y$ is the face of $T_y$ in the decomposition of $F$.

  	Now to see $\bar T_y \subseteq T_y$, we use a contradiction.
  	Suppose there was some $x \in \bar T_y$ such that $x \not \in T_y$.
  	Then there must be some $v_i \in V$ so that $\inprod {v_i}{x} > \sup_{y \in T_y} \inprod{v_i}{y}$.
  	This would imply that for all faces $F_{yj}$ of $T_y$, we have $v_i \not \in \N(F_{yj}, T_y)$.
  	However, by construction of $V$, we can consider the face $F$ of $T$ such that $v_i \in \N(F,T) = \cap_y \N(F_y, T_y)$ by Corollary~\ref{cor:face-decomp-normal-cones}.
  	We necessarily have $v_i \in \N(F_y, T_y)$, yielding a contradiction.
  	Thus, we have $v_k \in \N(F_y, T_y)$, so we conclude $\bar T_y \subseteq T_y$, yielding the equality of each $\bar T_y = T_y$.

  	Therefore, $V$ constructed as above is complete for both $T$ and each $T_y$ such that $T = \oplus_y T_y$.	
  	\jessie{Revisit and revise}
  \end{proof}

  \subsection{Once we have $V$}
  Now, for a given polytope $T(p) := \oplus p_y T_y$, we will later want to ask when a given $z \in \reals^d$ is in the polytope $T(p)$.
  We will then use this to understand the set of $p \in \simplex$ for which the given $z \in T(p)$.
  Throughout, assume we have $V$ which is complete for $\T$.

  Observe that as $T_y = \{x : Vx \leq E_{;y}\}$, we can multiply the right side of the inequality by the constant $p_y \geq  0$ to yield $p_y T_y = \{x : Vx \leq p_y E_{;y}\}$.
  Now in taking the Minkowski sum of polytopes described by the same set of normals, we can take 
  \begin{align*}
  \oplus_y p_y T_y &= \{x : Vx \leq p_1 E_{;1}\} \oplus \ldots \oplus \{x : Vx \leq p_n E_{;n}\} \\
  &= \{x : Vx \leq p_1 E_{;1} + \ldots + p_n E_{;n}\}\\
  &= \{x : Vx \leq E p\}~.~
  \end{align*}
  Observe the first to second line follows from Theorem~\ref{thm:support-minksum} and preservation of inequalities under addition.

  Now, we have $z \in T(p) \iff \inprod{v_i}{z} \leq (Ep)_i$ for all $v_i \in V$.
  As we are particularly interested in when $\vec 0 \in T(p)$, we have $\vec 0 \in T(p)$ if and only if $E p \geq 0$ by substitution.  \btw{For later, common result to say $T(p) = \partial \inprod p {L(\cdot)}$, so if $0 \in T(p)$, we have a minimizer of the loss.}

@@

  Still not sure if we can construct $B$ with the assumption that $p \in \simplex$ rather than $p \in A$, but gathering from our conversations at NeurIPS, that question seems worth investigating.
  I mean that in the sense that this $B$ matrix captures our necessary condition.
  The vertex representation seems to capture our sufficient condition, possibly rendering the simplex halfspace constraints unnecessary when we think of them as necessary conditions.

@@

 	Then $\{p \in \simplex : \vec 0 \in \oplus_y p_y T_y\} = \{p \in \simplex: Ep \geq \vec 0\}$.\btw{``stuff about $E$ and $D$.''}
  \end{lemma}
  \begin{proof}
    First, let us fix a distribution $p \in \simplex$.
    We define $T(p) := \oplus_y p_y T_y$ to be the $p$-weighted Minkowski sum over the polytopes in $\T$.
    By Theorem~\ref{thm:support-minksum}, we have the support of the (weighted) Minkowski sum is the (weighted) sum of the support of each polytope, which we can re-write as the product $Ep$.

@@

  \begin{definition}
  	We say a vector $v$ is \emph{redundant} with respect to matrix $Y$ describing the polytope $P^Y := \{z \in A : Yz \geq \vec 0\}$ if we have $P^Y = \{z \in A : [Y;v]z \geq \vec 0\}$.
  \end{definition}

  \begin{proposition}\label{prop:relate-rows}
  	Suppose we have polytopes $\T = \{T_1, \ldots, T_n\}$ and a set of normals $V$ that is complete for $\T$.
  	Take $E = (e_{iy})$ where $e_{iy} = \max_{x \in T_y} \inprod{v_i}{x}$, and take $D = \{p \in \simplex : Ep \geq \vec 0\}$ and $C = \{p \in A : Bp \geq \vec 0\}$.

  	Then $\{p \in \simplex : \vec 0 \in \oplus_y p_y T_y\} = C$	if and only if non-simplex rows of $B$ appear in $E$ (possibly scaled) and every other row of $E$ is redundant with respect to $B$.
  	\btw{``replace prop 1 statement with `non-simplex rows of $B$ appear in $E$ (possible scaled) and every other row of $E$ is redundant with respect to $B$.'''}
  \end{proposition}
  \begin{proof}
	\begin{itemize}
		\item [$\implies$] First, assume $C = \{p \in \simplex: \vec 0 \in \oplus_y p_y T_y\}$.
		By Proposition~\ref{prop:relate-E-B}, we know that $C = D := \{p \in \simplex : Ep \geq \vec 0\}$.
		We can re-write $D$ as $\{p \in A : [E;I_n] p \geq \vec 0\}$, building simplex constraints into the inequality rather than the domain assumption.
		By the assumed equality, every row of $B$ must then appear in $[E;I_n]$, where $I_n$ is the $n$-element identity matrix.
		As $\{p \in A : I_n p \geq \vec 0\}$ strictly defines the simplex, any non-simplex row of $B$ must then be described by a row of $E$, up to rescaling.

		What's more is that since the polytopes are the same, they have a unique minimal $\H$-representation (up to rescaling) since they are defined by an inequality on $\vec 0$.
		\btw{Elaborated this justification in the comment below ``As $D$ is defined... yielding a contradiction.''}
%		As $D$ is defined by distributions on $\simplex$ and $C$ on $A$, and $\simplex \subset A$, we have the rest of the rows in $E$ necessarily being redundant with respect to $B$.
%		To see this, suppose there was a row of $E$ that was not redundant with respect to $B$.
%		If such a row did not partition the simplex, it would necessarily be redundant by the restriction that $C \subseteq \simplex$.
%		Thus, such a row much partition the simplex, but not appear in $B$ up to rescaling.
%		Again, if the constraint does not partition the cell, the constraint is redundant, and if it does, then $C \neq D$, yielding a contradiction.

		\item [$\impliedby$] Suppose that any non-simplex rows of $B$ appear in $E$, and every other row of $E$ is redundant with respect to $B$.
		Then we have $D = \{p\in \simplex : Ep \geq \vec 0\} = \{p \in \simplex : Bp \geq \vec 0\}$.
		This polytope is equal to $\{p \in A : Bp \geq \vec 0\} = C$ since we require $C\subseteq \simplex$, so $B$ encapsulates any non-redundant simplex constraints.

@@

  \section{Necessary and sufficient conditions}
  \btw{This probably shouldn't be in this document... let's move it later.}
  \subsection{Necessary conditions}
    Halfspace representation
  \subsection{Sufficient conditions}
    Vertex representation
